---
  title: "Reproducible Research Assignment 1 - Week 2"
  output:
  html_document: default
  html_notebook: default
---

```{r global_options, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/Mike/Dropbox/MOOC/R Programming/RR/R_R_files/RepData_PeerAssessment1")
options(scipen = 999)
library(dplyr)
library(lattice)
```


#### Data Loading and Preprocessing :-

```{r}
setwd("C:/Users/Mike/Dropbox/MOOC/R Programming/RR/R_R_files/RepData_PeerAssessment1")
unzip('activity.zip')
Raw_Data <- read.csv('activity.csv')

```
#### Data Transform:-

```{r}
TempExtraDigit <- sprintf("%04d", Raw_Data$interval)
Raw_Data <- mutate(Raw_Data, interval = format(strptime(TempExtraDigit, format="%H%M"), format = "%H:%M"))
```


```{r}
DateSteps <- with(Raw_Data, tapply(steps, date, sum))
AverageDailySteps <- mean(DateSteps, na.rm = TRUE)
```

The average (mean) number of steps taken per day is `r AverageDailySteps`.


```{r}
TotalStepsPerDay_complete <- DateSteps[complete.cases(DateSteps)]
TotalStepsPerDay_complete
```

#### Histogram of the total number of steps taken each day:-

```{r, fig.width=8, fig.height=6}

hist(TotalStepsPerDay_complete, xlab = 'Daily Steps',
     main = "Histogram of Number of Steps Taken per Day")

```


The mean number of steps taken per day is `r round(mean(TotalStepsPerDay_complete),digits=1)` and the median value is `r round(median(TotalStepsPerDay_complete),digits=1)`.

#### Average daily activity pattern:-


```{r}
TimeSteps <- with(Raw_Data, tapply(steps, interval, mean, na.rm=TRUE))
plot(strptime(row.names(TimeSteps), format = "%H:%M"),TimeSteps,type="l",
     xlab = 'Time', ylab='Steps During 5-min. Interval',
main = "Time series plot of number of steps per day")
     
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

The 5-minute interval, on average across all the days in the dataset, which contained the maximum number of steps was at `r names(which(TimeSteps==max(TimeSteps)))`.

#### Input missing values:-


The number of rows with missing values is `r sum(complete.cases(Raw_Data) == FALSE)`.

Devise a strategy for filling in all of the missing values in the dataset: - 

New dataset that is equal to the original dataset but with the missing data filled in: -

```{r}
filled_data <- Raw_Data
filled_data[complete.cases(filled_data) == FALSE, ]$steps <- mean(filled_data$steps, na.rm=TRUE)
```

Histogram of the total number of steps taken each day: -

```{r, fig.width=8, fig.height=6}
FilledDateSteps <- with(filled_data, tapply(steps, date, sum))
```

The mean number of steps taken per day when the filled data is used is `r round(mean(TotalStepsPerDay_complete),digits=1)` compared to `r round(mean(TotalStepsPerDay_complete),digits=1)` for the unfilled data. The median value of the filled data was `r round(median(FilledDateSteps),digits=1)` compared to  `r round(median(TotalStepsPerDay_complete),digits=1)` for the unfilled data. The values are very similar and the missing data seems to have little impact on the results obtained.

```{r}

hist(FilledDateSteps, xlab = 'Daily Steps',
     main = "Histogram of Number of Steps Taken per Day (filled)")

```



#### Differences in activity patterns between weekdays and weekends: -


```{r}
filled_data <- mutate(filled_data, DayOfWeek = weekdays(as.Date(date)))
filled_data <- mutate(filled_data, DayType = factor(filled_data$DayOfWeek == 'Saturday' 
                           | filled_data$DayOfWeek == 'Sunday', 
                           labels = c('Weekday', 'Weekend')))
```



Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis): - 

```{r}
WeekendData <- filled_data[filled_data$DayType=='Weekend',]
Weekend <- with(WeekendData, tapply(steps, interval, mean, na.rm=TRUE))
WeekendFrame <- data.frame(Weekend)
WeekendFrame$DayType = as.factor('Weekend')
WeekendFrame <- rename(WeekendFrame, steps = Weekend)
WeekendFrame <- mutate(WeekendFrame, interval = row.names(WeekendFrame))

WeekdayData <- filled_data[filled_data$DayType=='Weekday',]
Weekday <- with(WeekdayData, tapply(steps, interval, mean, na.rm=TRUE))
WeekdayFrame <- data.frame(Weekday)
WeekdayFrame$DayType = as.factor('weekday')
WeekdayFrame <- rename(WeekdayFrame, steps = Weekday)
WeekdayFrame <- mutate(WeekdayFrame, interval = row.names(WeekdayFrame))

sorted <- rbind(WeekdayFrame, WeekendFrame)
```




```{r}
xyplot(steps~as.POSIXct(strptime(interval, format = "%H:%M"))|factor(DayType),
       sorted,
       scales=list(x=list(format="%H-%M")),
       type='l',
       layout=c(1,2),
       xlab='Time',ylab='Number of Steps')
````

It looks like people are more active in the at about 09:00 on weekdays compared to weekends.
knit2html()